Commutative character sheaves and geometric types for supercuspidal representations

### 2017-06-28 Geometric types.tex
Add section on character sheaves for SL(2,3), fix Section 3 to account for characters that don't vanish on Gder(k)

\usepackage{amssymb}
\usepackage{amsrefs}
% Fonts
\usepackage{mathrsfs}
% Enumitem
\usepackage{enumitem}
% Hyperrefs
\usepackage{hyperref}
\hypersetup{
  colorlinks   = true, %Colours links instead of ugly boxes
  urlcolor     = blue, %Colour for external hyperlinks
  linkcolor    = blue, %Colour of internal links
  citecolor   = green %Colour of citations
}
%\usepackage[margin=4.3cm]{geometry}
%\renewcommand{\baselinestretch}{1.0}

@@

      \newtheorem{conjecture}[theorem]{Conjecture}
%%%%%%%%%%%%%%% RINGS AND GROUPS %%%%%%%%%%%%%%%
\newcommand{\FF}{{\mathbb{F}}}
\newcommand{\ZZ}{{\mathbb{Z}}}
\newcommand{\NN}{{\mathbb{N}}}
\newcommand{\CC}{{\mathbb{C}}}
\newcommand{\QQ}{{\mathbb{Q}}}
\newcommand{\RR}{{\mathbb{R}}}
\newcommand{\EE}{\mathbb{\bar Q}_\ell}
\newcommand{\OK}{\mathcal{O}_K}
\newcommand{\OL}{\mathcal{O}_L}
\newcommand{\OO}[1]{\mathcal{O}_{#1}}
\newcommand{\bFq}{\bar{k}}
\newcommand{\Fq}{k}
\newcommand{\Fqm}{k_m}
\newcommand{\EEx}{\EE^\times}
\newcommand{\ZEx}{\mathbb{\bar Z}_\ell^\times}
\newcommand{\Weil}[1]{\mathcal{W}_{#1}}
\newcommand{\m}{{\mathfrak{m}}}
%%%%%%%%%%%%%%% ALGEBRAIC GROUPS %%%%%%%%%%%%%%%
\newcommand{\Gm}[1]{\mathbb{G}_{\hskip-1pt\textbf{m},#1}}
\DeclareMathOperator{\GL}{GL}
\newcommand{\comp}{\Pi} % Component group
\newcommand{\G}{\textbf{G}}
%%%%%%%%%%%%%%% NAMED OPERATORS %%%%%%%%%%%%%%%
\DeclareMathOperator{\Gal}{Gal}
\newcommand{\Frob}[1]{\operatorname{Fr}_{#1}}
\DeclareMathOperator{\Aut}{Aut}
\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\ord}{ord}
\DeclareMathOperator{\coker}{coker}
\DeclareMathOperator{\Gr}{Gr}
\DeclareMathOperator{\Irrep}{Irrep}
\DeclareMathOperator{\id}{id}
\DeclareMathOperator{\Ext}{Ext}
\DeclareMathOperator{\Hh}{H}
\DeclareMathOperator{\Res}{Res}
\DeclareMathOperator{\Nm}{Nm}
\DeclareMathOperator{\trace}{Tr}
\DeclareMathOperator{\obj}{obj}
\DeclareMathOperator{\mor}{mor}
\DeclareMathOperator{\Lang}{Lang}
\DeclareMathOperator{\image}{im}
\DeclareMathOperator{\Loc}{Loc}
\DeclareMathOperator{\Tot}{Tot}
\DeclareMathOperator{\Tor}{Tor}
\DeclareMathOperator{\SL}{SL}
\DeclareMathOperator{\PGL}{PGL}
\newcommand{\gal}[1]{{\operatorname{Gal}\hskip-1pt\left( {\bar #1}/#1 \right)}}
\newcommand{\Spec}[1]{{\operatorname{Spec}(#1)}}
\newcommand{\op}{_{\operatorname{op}}}
\newcommand{\der}{_{\operatorname{der}}}
\newcommand{\ab}{_{\operatorname{ab}}}

%%%%%%%%%%%% MISCELLANEOUS OPERATORS %%%%%%%%%%%%
\newcommand{\sheafHom}{{\mathscr{H}\hskip-4pt{\it o}\hskip-2pt{\it m}}}
\newcommand{\abs}[1]{{\vert #1 \vert}}
\newcommand{\ceq}{{\, :=\, }}
\newcommand{\tq}{{\ \vert\ }}
\newcommand{\iso}{{\ \cong\ }}
\newcommand{\trFrob}[1]{t_{#1}}
\DeclareMathOperator{\Tr}{Tr}
\newcommand{\TrFrob}[1]{\Tr_{#1}}
%% Limits
\newcommand{\invlim}[1]{\lim\limits_{\overleftarrow{#1}}}
\newcommand{\dirlim}[1]{\lim\limits_{\overrightarrow{#1}}}
\newcommand{\limit}[1]{\mathop{\textsc{lim}}\limits_{#1}}
\newcommand{\colimit}[1]{\mathop{\textsc{colim}}\limits_{#1}}
%% Fonts for quasicharacter sheaves
\newcommand{\cs}[1]{{\mathcal{#1}}}
\newcommand{\gcs}[1]{{\mathcal{\bar #1}}}
\newcommand{\dualgcs}[1]{\gcs{#1}^\dagger}
\newcommand{\dualcs}[1]{\cs{#1}^\dagger}
%% Categories
\newcommand{\CS}{{\mathcal{C\hskip-0.8pt S}}}
\newcommand{\CCS}{{\mathcal{C\hskip-.8pt C\hskip-0.8pt S}}}
\newcommand{\bCS}{{\CS_0}}
\newcommand{\CSab}{\CS_{\operatorname{ab}}}
\newcommand{\catname}[1]{\normalfont{\textsf{#1}}}
\newcommand{\Sch}[1]{{\catname{Sch}_{/#1}}}
\newcommand{\QCS}{{\mathcal{QC\hskip-0.8pt S}}}
\newcommand{\CSiso}[1]{\CS(#1)_{/\text{iso}}}
\newcommand{\bCSiso}[1]{\bCS(#1)_{/\text{iso}}}
\newcommand{\QCSiso}[1]{\QCS(#1)_{/\text{iso}}}
\newcommand{\CCSiso}[1]{\CCS(#1)_{/\text{iso}}}
%% Labeled items
\makeatletter
\newcommand{\labitem}[2]{
\def\@itemlabel{\textbf{#1}}
\item
\def\@currentlabel{#1}\label{#2}}
\makeatother
%% Shorthand for bars
\renewcommand{\bf}{\bar{f}}
\newcommand{\bg}{{\bar{g}}}
\newcommand{\bm}{\bar{m}}
\newcommand{\bG}{\bar{G}}
\newcommand{\bH}{\bar{H}}
\newcommand{\brho}{{\bar\rho}}
\newcommand{\bx}{{\bar{x}}}
%% Spacing control
\newcommand{\tight}[3]{\hspace{-#1pt}{#2}\hspace{-#3pt}}
\newcommand{\GxG}{\text{$G \tight{1}{\times}{1} G$}}
\newcommand{\bGxG}{\text{$\bar{G} \tight{1}{\times}{1} \bar{G}$}}
\newcommand{\bfxf}{\text{$\bar{f} \tight{1}{\times}{1} \bar{f}$}}
\newcommand{\GxxG}{\text{$G \tight{1}{\times}{1} G$}}
\newcommand{\LxL}{\text{$\gcs{L} \tight{0}{\boxtimes}{0} \gcs{L}$}}

@@

@@

We now consider the case of a smooth group scheme without the commutativity assumption.  We start
by relating linear character sheaves on $G$ to character sheaves on its abelianization.

If $\chi \in G(k)^*$ is a character, it must vanish on the derived subgroup $G\der(k)$. \todo{Error} 
Kamgarpour gives an example \cite{kamgarpour:09a}*{(1.1)} of a linear character sheaf
on $G = \PGL_n$ that does not vanish on $G\der = \PGL_n$, defined by the extension
\[
1 \to \mu_n \to \SL_n \to \PGL_n \to 1
\]
together with a non-trivial character of $\mu_n(k)$. \todo{Small change here.}
In order to obtain a relationship between linear character sheaves on $G$ and characters of $G(k)$,
he opts to give a different definition of commutator and, in doing so, introduces a `stacky abelianization' of $G$ in order to geometrize characters of $G(\Fq)$.
 Since we have already seen the need to adapt the
notion of character sheaf, even in the commutative case, we instead add restrictions
to the definition of commutative character sheaf and leave the definition of $G\der$ unchanged, allowing us to use the schematic abelianization of $G$ in the geometrization of characters of $G(\Fq)$.

We begin this section with the main definition in this paper - the category $\CCS(G)$ of commutative character sheaves, Definition~\ref{def:CCS}. 
This definition is delicate and somewhat technical, but it is vindicated in Theorem~\ref{thm:Gab} which shows that $\CCS(G)$ is equivalent to the category of commutative character sheaves on the abelianization $G\ab$ of $G$. 
To prove Theorem~\ref{thm:Gab} we use descent theory in Section~\ref{ssec:descent}, in the process giving insight into Definition~\ref{def:CCS}. 
Section~\ref{sec:noncom} concludes with Theorem~\ref{thm:geo}, showing that the dictionary from $\CCS(G)$ to characters of $G(k)$ is as precise as possible.

@@

Theorem~\ref{thm:Gab} shows that $\CCS(G)$ is a categorical solution to the problem that linear character sheaves on $G$ need not be trivial on $G\der$, as discussed at the beginning of Section~\ref{sec:noncom}.  At the same time, changing $\CS(G)$ to $\CCS(G)$ resolves the lack of bijectivity in \cite{cunningham-roe:13a}*{Theorem 3.6}.
We may also use Theorem~\ref{thm:Gab} to give a description of the morphisms and the isomorphism classes of objects in $\CCS(G)$.

@@

\begin{theorem}\label{thm:geo}
The trace of Frobenius $\TrFrob{} : \CCSiso{G}\to G(\Fq)^*$ fits into the following diagram,
\[
\begin{tikzcd}
\ & & \CCSiso{G\ab} \arrow{d}{\TrFrob{}}[swap]{\iso} \arrow{r}{\iso} & \CCSiso{G} \arrow{d}{\TrFrob{}}& \\
1 \arrow{r} & \Delta_G^* \arrow{r} & G\ab(k)^* \arrow{r} & G(k)^* \arrow{r} & 1,
\end{tikzcd}
\]
where $\Delta_G$ is the image of the connecting homomorphism $G\ab(\Fq) \to \Hh^1(\Fq, G\der)$. 
Thus the category $\CCS(G)$ geometrizes characters of $G(\Fq)$ in the following sense: for every group homomorphism $\chi : G(\Fq) \to \EEx$ there is an object $(\cs{L},\mu,\beta)$ in $\CCS(G)$ such that $\trFrob{\cs{L}} = \chi$. 
While the geometrization of $\chi : G(\Fq) \to \EEx$ is not unique, the group of isomorphism classes of possibilities are enumerated by $\Delta_G^*$.
\end{theorem}

\begin{proof}
By the definition of $\Delta_G$, we have a short exact sequence
\[
1 \to G(k) / G\der(k) \to G\ab(k) \to \Delta_G \to 1.
\]
Applying $\Hom(-, \EEx)$ and using the fact that every homomorphism $G(k) \to \EEx$ vanishes on $G\der(k)$,\todo{Error here} we get
\[
1 \to \Delta_G^* \to G\ab(k)^* \to G(k)^* \to 1.
\]
By Theorem~\ref{thm:Gab}, the map $\CCSiso{G\ab} \to \CCSiso{G}$ is an isomorphism.
Moreover, since both $\CCSiso{G\ab} \to \CCSiso{G}$ and $G\ab(k)^* \to G(k)^*$ are
defined by pullback along $q$, the square in the statement of the theorem commutes.
Finally, $\TrFrob{} : \CCSiso{G\ab} \to G\ab(k)^*$ is an isomorphism by Corollary \ref{cor:Gab}.
\end{proof}

\begin{remark}
Note that when $\Hh^1(k,G\der) = 0$ then $\CCSiso{G} \iso G(\Fq)^*$, so
we succeed in geometrizing characters of $G(k)$ on the nose.
\end{remark}

@@

>>>>

\usepackage{amssymb}
\usepackage{amsrefs}
\usepackage{multirow}
% Fonts
\usepackage{mathrsfs}
% Enumitem
\usepackage{enumitem}
% Hyperrefs
\usepackage{hyperref}
\hypersetup{
  colorlinks   = true, %Colours links instead of ugly boxes
  urlcolor     = blue, %Colour for external hyperlinks
  linkcolor    = blue, %Colour of internal links
  citecolor   = green %Colour of citations
}
%\usepackage[margin=4.3cm]{geometry}
%\renewcommand{\baselinestretch}{1.0}

@@

      \newtheorem{conjecture}[theorem]{Conjecture}
%%%%%%%%%%%%%%% RINGS AND GROUPS %%%%%%%%%%%%%%%
\newcommand{\FF}{{\mathbb{F}}}
\newcommand{\FFF}{{\mathbf{F}_3}}
\newcommand{\ZZ}{{\mathbb{Z}}}
\newcommand{\NN}{{\mathbb{N}}}
\newcommand{\CC}{{\mathbb{C}}}
\newcommand{\QQ}{{\mathbb{Q}}}
\newcommand{\RR}{{\mathbb{R}}}
\newcommand{\EE}{\mathbb{\bar Q}_\ell}
\newcommand{\OK}{\mathcal{O}_K}
\newcommand{\OL}{\mathcal{O}_L}
\newcommand{\OO}[1]{\mathcal{O}_{#1}}
\newcommand{\bFq}{\bar{k}}
\newcommand{\Fq}{k}
\newcommand{\Fqm}{k_m}
\newcommand{\EEx}{\EE^\times}
\newcommand{\ZEx}{\mathbb{\bar Z}_\ell^\times}
\newcommand{\Weil}[1]{\mathcal{W}_{#1}}
\newcommand{\m}{{\mathfrak{m}}}
%%%%%%%%%%%%%%% ALGEBRAIC GROUPS %%%%%%%%%%%%%%%
\newcommand{\Gm}[1]{\mathbb{G}_{\hskip-1pt\textbf{m},#1}}
\DeclareMathOperator{\GL}{GL}
\newcommand{\comp}{\Pi} % Component group
\newcommand{\G}{\textbf{G}}
%%%%%%%%%%%%%%% NAMED OPERATORS %%%%%%%%%%%%%%%
\DeclareMathOperator{\Gal}{Gal}
\newcommand{\Frob}[1]{\operatorname{Fr}_{#1}}
\DeclareMathOperator{\Aut}{Aut}
\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\ord}{ord}
\DeclareMathOperator{\coker}{coker}
\DeclareMathOperator{\Gr}{Gr}
\DeclareMathOperator{\Irrep}{Irrep}
\DeclareMathOperator{\id}{id}
\DeclareMathOperator{\Ext}{Ext}
\DeclareMathOperator{\Hh}{H}
\DeclareMathOperator{\Res}{Res}
\DeclareMathOperator{\Nm}{Nm}
\DeclareMathOperator{\trace}{Tr}
\DeclareMathOperator{\obj}{obj}
\DeclareMathOperator{\mor}{mor}
\DeclareMathOperator{\Lang}{Lang}
\DeclareMathOperator{\image}{im}
\DeclareMathOperator{\Loc}{Loc}
\DeclareMathOperator{\Tot}{Tot}
\DeclareMathOperator{\Tor}{Tor}
\DeclareMathOperator{\SL}{SL}
\DeclareMathOperator{\PGL}{PGL}
\DeclareMathOperator{\PSL}{PSL}
\newcommand{\gal}[1]{{\operatorname{Gal}\hskip-1pt\left( {\bar #1}/#1 \right)}}
\newcommand{\Spec}[1]{{\operatorname{Spec}(#1)}}
\newcommand{\op}{_{\operatorname{op}}}
\newcommand{\der}{_{\operatorname{der}}}
\newcommand{\ab}{_{\operatorname{ab}}}

%%%%%%%%%%%% MISCELLANEOUS OPERATORS %%%%%%%%%%%%
\newcommand{\sheafHom}{{\mathscr{H}\hskip-4pt{\it o}\hskip-2pt{\it m}}}
\newcommand{\abs}[1]{{\vert #1 \vert}}
\newcommand{\ceq}{{\, :=\, }}
\newcommand{\tq}{{\ \vert\ }}
\newcommand{\iso}{{\ \cong\ }}
\newcommand{\trFrob}[1]{t_{#1}}
\DeclareMathOperator{\Tr}{Tr}
\newcommand{\TrFrob}[1]{\Tr_{#1}}
%% Limits
\newcommand{\invlim}[1]{\lim\limits_{\overleftarrow{#1}}}
\newcommand{\dirlim}[1]{\lim\limits_{\overrightarrow{#1}}}
\newcommand{\limit}[1]{\mathop{\textsc{lim}}\limits_{#1}}
\newcommand{\colimit}[1]{\mathop{\textsc{colim}}\limits_{#1}}
%% Fonts for quasicharacter sheaves
\newcommand{\cs}[1]{{\mathcal{#1}}}
\newcommand{\gcs}[1]{{\mathcal{\bar #1}}}
\newcommand{\dualgcs}[1]{\gcs{#1}^\dagger}
\newcommand{\dualcs}[1]{\cs{#1}^\dagger}
\newcommand{\IC}{\mathcal{IC}}
%% Categories
\newcommand{\CS}{{\mathcal{C\hskip-0.8pt S}}}
\newcommand{\CCS}{{\mathcal{C\hskip-.8pt C\hskip-0.8pt S}}}
\newcommand{\bCS}{{\CS_0}}
\newcommand{\CSab}{\CS_{\operatorname{ab}}}
\newcommand{\catname}[1]{\normalfont{\textsf{#1}}}
\newcommand{\Sch}[1]{{\catname{Sch}_{/#1}}}
\newcommand{\QCS}{{\mathcal{QC\hskip-0.8pt S}}}
\newcommand{\CSiso}[1]{\CS(#1)_{/\text{iso}}}
\newcommand{\bCSiso}[1]{\bCS(#1)_{/\text{iso}}}
\newcommand{\QCSiso}[1]{\QCS(#1)_{/\text{iso}}}
\newcommand{\CCSiso}[1]{\CCS(#1)_{/\text{iso}}}
%% Labeled items
\makeatletter
\newcommand{\labitem}[2]{
\def\@itemlabel{\textbf{#1}}
\item
\def\@currentlabel{#1}\label{#2}}
\makeatother
%% Shorthand for bars
\renewcommand{\bf}{\bar{f}}
\newcommand{\bg}{{\bar{g}}}
\newcommand{\bm}{\bar{m}}
\newcommand{\bG}{\bar{G}}
\newcommand{\bH}{\bar{H}}
\newcommand{\brho}{{\bar\rho}}
\newcommand{\bx}{{\bar{x}}}
%% Spacing control
\newcommand{\tight}[3]{\hspace{-#1pt}{#2}\hspace{-#3pt}}
\newcommand{\GxG}{\text{$G \tight{1}{\times}{1} G$}}
\newcommand{\bGxG}{\text{$\bar{G} \tight{1}{\times}{1} \bar{G}$}}
\newcommand{\bfxf}{\text{$\bar{f} \tight{1}{\times}{1} \bar{f}$}}
\newcommand{\GxxG}{\text{$G \tight{1}{\times}{1} G$}}
\newcommand{\LxL}{\text{$\gcs{L} \tight{0}{\boxtimes}{0} \gcs{L}$}}

@@

% Matrices
\newcommand{\smat}[4]{$\left(\begin{smallmatrix}#1 & #2 \\ #3 & #4 \end{smallmatrix}\right)$}

@@

We now consider the case of a smooth group scheme without the commutativity assumption.  Our approach
is to relate linear character sheaves on $G$ to character sheaves on its abelianization $G\ab = G / G\der$.
While every character $\chi \in G(k)^*$ vanishes on $G(k)\der$, it may not vanish on $G\der(k)$.  For example,
if $k$ has odd characteristic then there are nontrivial characters $\PGL_2(k) \to \EEx$ vanishing on
$\PGL_2(k)\der = \PSL_2(k)$ (see Section \ref{sec:badchar}).  In passing to $G\ab$,
we may only hope to geometrize characters that vanish on all of $G\der(k)$.

We begin this section with the main definition in this paper - the category $\CCS(G)$ of commutative character sheaves, Definition~\ref{def:CCS}. 
This definition is delicate and somewhat technical, but it is vindicated in Theorem~\ref{thm:Gab} which shows that $\CCS(G)$ is equivalent to the category of commutative character sheaves on the abelianization $G\ab$ of $G$. 
To prove Theorem~\ref{thm:Gab} we use descent theory in Section~\ref{ssec:descent}, in the process giving insight into Definition~\ref{def:CCS}. 
Section~\ref{sec:noncom} concludes with Theorem~\ref{thm:geo}, showing that the dictionary from $\CCS(G)$
to characters of $G(k)$ in fact encompasses every character vanishing on $G\der(k)$.

@@

We may use Theorem~\ref{thm:Gab} to give a description of the morphisms and the isomorphism classes of objects in $\CCS(G)$.

@@

\begin{theorem}\label{thm:geo}
The trace of Frobenius $\TrFrob{} : \CCSiso{G}\to G(\Fq)^*$ fits into the following diagram,
\[
\begin{tikzcd}
\ & & \CCSiso{G\ab} \arrow{d}{\TrFrob{}}[swap]{\iso} \arrow{r}{\iso} & \CCSiso{G} \arrow{d}{\TrFrob{}}& \\
1 \arrow{r} & \Delta_G^* \arrow{r} & G\ab(k)^* \arrow{r} & G(k)^* \arrow{r} & G\der(k)^* \arrow{r} & 1,
\end{tikzcd}
\]
where $\Delta_G$ is the image of the connecting homomorphism $G\ab(\Fq) \to \Hh^1(\Fq, G\der)$. 
Thus the category $\CCS(G)$ geometrizes characters of $G(\Fq)$ in the following sense: for every group homomorphism $\chi : G(\Fq) \to \EEx$ that vanishes on $G\der(\Fq)$, there is an object $(\cs{L},\mu,\beta)$ in $\CCS(G)$ such that $\trFrob{\cs{L}} = \chi$. 
While the geometrization of $\chi$ is not unique, the isomorphism classes of possibilities are enumerated by $\Delta_G^*$.
\end{theorem}

\begin{proof}
By the definition of $\Delta_G$, we have a short exact sequence
\[
1 \to G\der(k) \to G(k) \to G\ab(k) \to \Delta_G \to 1.
\]
Applying $\Hom(-, \EEx)$ yields the bottom row in the statement.

By Theorem~\ref{thm:Gab}, the map $\CCSiso{G\ab} \to \CCSiso{G}$ is an isomorphism.
Moreover, since both $\CCSiso{G\ab} \to \CCSiso{G}$ and $G\ab(k)^* \to G(k)^*$ are
defined by pullback along $q$, the square in the statement of the theorem commutes.
Finally, $\TrFrob{} : \CCSiso{G\ab} \to G\ab(k)^*$ is an isomorphism by Corollary \ref{cor:Gab}.
\end{proof}

\section{Examples of Nontrivial Characters on the Derived Subgroup} \label{sec:badchar}

In this section we give two examples of characters of $G(k)$ that do not vanish on $G\der(k)$.

\subsection{\texorpdfstring{$\PGL_2(k)$}{PGL(2, k)}}

Let $G = \PGL_2$ over a finite field $k$ of odd characteristic.  Then $G\der = G$, but $G(k)\der = \PSL_2(k)$ has index $2$ in $G(k)$.  There is thus a nontrivial character $\chi : G(k) \to \{\pm 1\} \subset \EEx$.  We may construct a linear character sheaf with trace $\chi$ as follows.  Consider the standard isogeny $f : \SL_2 \to \PGL_2$ with kernel $\mu_2$ and the nontrivial map $\psi : \mu_2 \to \EEx$.  Then the pushforward $f_! \EE$ of the constant sheaf on $\SL_2$ is a $2$-dimensional local system, and its $\psi$-isotypic component $(f_! \EE)_\psi$ is a linear character sheaf on $\PGL_2$ with trace of Frobenius $\chi$ (c.f. \cite{cunningham-roe:13a}*{Prop 1.10}).

\subsection{\texorpdfstring{$\SL_2(\FFF)$}{\SL(2, 3)}}

Let $G = \SL_2$ over $\FFF$.  Then $G\der = G$, but $G(k)\der$ is the subgroup of semisimple elements, isomorphic to the quaternion group $Q_8$ and of index $3$.  There is thus a nontrivial character $\chi : G(k) \to \mu_3 \subset \EEx$.  The method described above for producing a linear character sheaf corresponding to this character no longer works, since $\SL_2$ is simply connected and thus has no discrete isogenies mapping onto it.  We may still produce a local system with the correct trace of Frobenius, by pushing forward the constant sheaf along the Lang isogeny and picking out the isotypic component corresponding to $\chi$.  However, since $\SL_2$ is nonabelian, the Lang isogeny is not a homomorphism and the resulting local system is not a linear character sheaf.

Alternatively, we can use the fact that Lusztig's character sheaves provide a basis for the space of class functions \cite{lusztig:86a}*{Cor. 25.7} to give a different geometrization of $\chi$.  Lusztig's character sheaves are constructed by perverse extension of local systems on a certain stratification of $G$.  In the case of $\SL_2$, there are five strata:

\vspace{0.1in}

\begin{tabular}{|l|l|}
\hline
$C_3$ & regular semisimple elements \\
$C_2$ & regular unipotent elements \\
$C_2^-$ & $\{-u : \mbox{ regular unipotent $u$}\}$ \\
$1$ & $\{1\}$ \\
$-1$ & $\{-1\}$ \\
\hline
\end{tabular}

\vspace{0.1in}

Suppose that $C$ is a stratum and $x \in C$. Local systems on $C$ are classified up to isomorphism by irreducible representations of $\pi_0(Z_G(x))$.  This component group is trivial for $C_3$, $1$ and $-1$, and of order $2$ for $C_2$ and $C_2^-$.  Denote the nontrivial local systems on $C_2$ and $C_2^-$ by $\cs{E}_{C_2}$ and $\cs{E}_{C_2^-}$.  Then the character sheaves on $\SL_2$ are the following, where the second column gives a decomposition in the Grothendieck group of standard sheaves on $G$.

\vspace{0.1in}

\begin{tabular}{|l|l|ccccccc|}
\hline
Perverse & Standard & \multicolumn{7}{|c|}{Trace of Frobenius values} \\ \cline{3-9}

Sheaf & Sheaves & \smat1001 & \smat2002 & \smat0120 & \smat1101 & \smat1201 & \smat2102 & \smat2202 \\
\hline
$\IC(\mathbf{1}_{C_3})$ & $\mathbf{1}_{C_3}[3]$ & -1 & -1 & -1 & -1 & -1 & -1 & -1 \\
$\IC(\mathbf{1}_{C_2})$ & $\mathbf{1}_{C_2}[2] \oplus \mathbf{1}_1[0]$ & 1 & 0 & 0 & 1 & 1 & 0 & 0 \\
$\IC(\cs{E}_{C_2})$ & $\cs{E}_{C_2}[2]$ & 0 & 0 & 0 & $\sqrt{-3}$ & $-\sqrt{-3}$ & 0 & 0 \\
$\IC(\mathbf{1}_{C_2^-})$ & $\mathbf{1}_{C_2^-}[2] \oplus \mathbf{1}_{-1}[0]$ & 0 & 1 & 0 & 0 & 0 & 1 & 1 \\
$\IC(\cs{E}_{C_2^-})$ & $\cs{E}_{C_2^-}[2]$ & 0 & 0 & 0 & 0 & 0 & $\sqrt{-3}$ & $-\sqrt{-3}$ \\
$\IC(\mathbf{1}_1)$ & $\mathbf{1}_1[0]$ & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
$\IC(\mathbf{1}_{-1})$ & $\mathbf{1}_{-1}[0]$ & 0 & 1 & 0 & 0 & 0 & 0 & 0 \\
\hline
\end{tabular}

\vspace{0.1in}
\noindent Using the trace of Frobenius values, we find that 
\[
- \IC(\mathbf{1}_{C_3}) - \frac32 \IC(\mathbf{1}_{C_2}) + \frac12 \IC(\cs{E}_{C_2}) - \frac32 \IC(\mathbf{1}_{C_2^-}) - \frac12 \IC(\cs{E}_{C_2^-}) + \frac32 \IC(\mathbf{1}_1) + \frac32 \IC(\mathbf{1}_{-1})
\]
also geometrizes $\chi$.  Note that all seven character sheaves appear in this geometrization, even though we start just with a character.

\subsection{The True Commutator}

Kamgarpour gives an example \cite{kamgarpour:09a}*{(1.1)} of a linear character sheaf
on $G = \PGL_n$ that does not vanish on $G\der = \PGL_n$, defined by the extension
\[
1 \to \mu_n \to \SL_n \to \PGL_n \to 1
\]
together with a non-trivial character of $\mu_n(k)$. \todo{Small change here.}
In order to obtain a relationship between linear character sheaves on $G$ and characters of $G(k)$,
he opts to give a different definition of commutator and, in doing so, introduces a `stacky abelianization' of $G$ in order to geometrize characters of $G(\Fq)$.
 Since we have already seen the need to adapt the
notion of character sheaf, even in the commutative case, we instead add restrictions
to the definition of commutative character sheaf and leave the definition of $G\der$ unchanged, allowing us to use the schematic abelianization of $G$ in the geometrization of characters of $G(\Fq)$.

@@

\bib{lusztig:86a}{article}{
	Author = {George Lusztig},
	Date-Added = {2017-06-28 19:26:19 +0000},
	Date-Modified = {2017-06-28 19:28:23 +0000},
	Journal = {Advances in Math.},
	Keywords = {qcsheaves},
	Number = {2},
	Pages = {{103--155}},
	Title = {{Character sheaves V}},
	Volume = {61},
	Year = {1986}}
